---
title: "HMP 669 project contribution"
output: html_notebook
---

```{r}
library(tidylab)
library(tidyverse)
library(ggplot2)
library(skimr)
```

A suggested (non-exhaustive) list of financial measures includes:

Profitability of patients from Cardiac and Orthopedic services
LOS outlier rates
Payer mix
Rates of emergent vs. elective care
Probable future demand given Omega's patient demographics for Cardiac and Orthopedic services (this is less of a coding/query question, and more of discussing what you can an infer from demographics and trends)

```{r}
finance = read_csv("OMEGA_2000.csv")
service = read_csv("SERVICE.csv")
```

#filter by Physicians, specialized in Ortho and 
```{r}
finance1 = 
  finance %>% 
  group_by(DRG) %>% 
  filter(DRG == 5 | DRG == 8)
```


#Selected on DRG
```{r}
  finance1 %>%
  mutate(DRG = if_else(DRG == 5, "cardio", "ortho")) %>% 
  ggplot(aes(x = DRG, y = LOS, fill = DRG)) +
  geom_violin() 
```

```{r}
finance1 %>%
  mutate(DRG = if_else(DRG == 5, "cardio", "ortho")) %>% 
  ggplot(aes(x = DRG, y = LOS, fill = DRG)) +
  geom_boxplot() 
```

##Selected on PHY_A_SPC
```{r}
finance2 = 
  finance %>% 
  group_by(PHYS_A_SPC) %>% 
  filter(PHYS_A_SPC == 14 | PHYS_A_SPC == 58)
```

```{r}
finance2 %>%
  mutate(PHYS_A_SPC = if_else(PHYS_A_SPC == 14, "cardio", "ortho")) %>% 
  ggplot(aes(x = PHYS_A_SPC, y = LOS, fill = PHYS_A_SPC)) +
  geom_boxplot() 
```
```{r}
finance1 %>%
  mutate(DRG = if_else(DRG == 5, "cardio", "ortho")) %>% 
  ggplot(aes(x = DRG, y = LOS, fill = DRG)) +
  geom_boxplot()
```


```{r}
  group_by(PHYS_A_SPC) %>% 
  mutate(payment_prop = sum(TOT_PAY)/total_payment)
```

```{r}
finance41 = 
  finance2 %>% 
  group_by(PHYS_A_SPC) %>% 
  summarize(total_pay = sum(TOT_PAY))
```
 
```{r}
finance41 %>%
  mutate(PHYS_A_SPC = if_else(PHYS_A_SPC == 14, "cardio", "ortho")) %>% 
  ggplot(aes(x = PHYS_A_SPC, y = total_pay)) +
  geom_bar()
```

```{r}
fin_service = 
  full_join(finance, service, by = c("PHYS_A_SPC" = "SERVICE"))
```
 
```{r}
head(fin_service)
```
 
            
```{r}
skim(fin_service)
```


##1
#compare with national benchmarks, the whole hop versus other service lines

A. Revenue

```{r}
    fin_service %>% 
    group_by(DESCRIPT) %>% 
    dplyr::summarize(total_pay = sum((TOT_PAY), na.rm = TRUE)) %>% 
    dplyr::mutate(payment_prop = ((total_pay/sum(total_pay, na.rm = TRUE)*100))) %>%
    filter(payment_prop > 3) %>% 
    ggplot(aes(x = reorder(DESCRIPT, - payment_prop), y = payment_prop)) +
    geom_bar(stat = 'identity')
```

B. Profit

```{r}
service_profit = 
  fin_service %>% 
  group_by(PHYS_A_SPC) %>% 
  dplyr::mutate(profitability = (TOT_PAY - TOT_COST)/TOT_PAY, na.rm = T) %>% 
  select(DESCRIPT, PHYS_A_SPC, profitability) %>% 
  arrange(desc(profitability))
  


service_profit
``` 


#Comparing costs
```{r}
finance_payment %>% 
  filter(PHYS_A_SPC == 14 | PHYS_A_SPC == 18) %>% 
  mutate(PHYS_A_SPC = if_else(PHYS_A_SPC == 14, "cardio", "ortho" )) %>% 
  ggplot(aes(PHYS_A_SPC))
```


##2 Box Plot

##3 Medicaid pays less (low income patients coming to the hospital), compare uninsured, payment, missing, etc.
How tht total income varies amongst payer


##4 emergent: why they picked that particular excel.